Kratos: A solution for data privacy,\ literacy and student agency in a data-driven school ecosystem

### 2019-06-07 paper.tex
add submission draft

\textit{Abstract}
Growing digitization has made data ownership an important focus point for institutions and students.\ Broadly,\ there are three issues,\ which require urgent attention for optimization of data privacy,\ literacy,\ and utilization.\ First,\ schools globally are equivocal about data generated by and about students as a result of the digitization of instruction,\ learning,\ and assessment.\ They lack necessary frameworks for data literacy,\ data interoperability,\ and optimization,\ while maintaining privacy and control.\ Second,\ the scale,\ source,\ and nature of school data makes its interoperability impractical,\ resulting in an inability to assess the true impact of educational technologies on instruction and learning.\ Third,\ while data helps teachers improve pedagogical practices,\ an increasingly data-driven decision-making process suggests that student dimensions of learning and equitable participation in curriculum design becomes secondary.\ Finding a balance between data-driven decision making and student voice is critical for an efficient school ecosystem.\ In this paper,\ we introduce Kratos: an immutable decentralised data management system that provides data privacy and applied data literacy while empowering students with a user interface for data governance and active participation in the school ecosystem.\ Using the advantages of blockchain technologies,\ Kratos enables easy authentication and access to data.\ The objective of Kratos is thus to equip schools and students with the ability to access,\ manage and control their data and to understand how,\ why,\ and by whom data is accessed without compromising student agency and privacy.\ This paper describes proof of concept for Kratos and its benefits to the school ecosystem.
\bigbreak
\newpage
\section{Introduction}

To improve work,\ school practitioners need data interoperability and educational information from data mining (Bienkowski,\ Feng,\ and Means,\ 2012; Snyder et al.,\ 2016).\ This need emanates from the requirement to measure school effectiveness by combining student achievement,\ data,\ and accountability (Militello et al.,\ 2013).

@@

\subsection{Data control}
In conventional systems,\ data control or 'ownership' can be proved with the help of an access token but this provides no guarantee on when the owner came into possession of the data.\ In order to attest 'ownership' at a specific point in time,\ we need time-stamping services like those described by Todd (2016),\ Szalachowski (2018),\ and Massias et al.(1999).\ A time-stamping service requires something to be committed along with the time-stamp and Kratos envisions this to be a cryptographic hash which also acts as an access token which vendors can use to access any student-generated data.
\bigbreak
The aim of using a cryptographic hash like SHA-3 (Bertoni et al.,\ 2012) is to ensure a uniquely random reference to the data to avoid out-of-channel data leaks.\ Kratos suggests using an element of randomness like a salt to generate the hash in order to have the ability to revoke tokens by regenerating randomness.\ Kratos enforces that all schools encrypt their data before creating access tokens to mitigate the risk of loss/theft of data.\ Past research (Zetter,\ 2016) shows that firms are willing to circumnavigate laws to collect data and encrypting data by design ensures that no third party can have access without being granted so explicitly.
\bigbreak
Kratos suggests that encryption be done at the student level but this can also be at the school level depending on existing frameworks and rules surrounding the school.\ If the school chooses to encrypt student data on behalf of the student,\ Kratos enforces that the school use a unique key for each individual to minimize the risk of key theft.\ In addition to this,\ Kratos suggests that schools store their encrypted data in a distributed file storage system like IPFS to ensure data redundancy and availability in case of a setback.\ Storing data on IPFS also makes it easier to create time-stamps since it is sufficient to reference the IPFS pointer instead of potentially hashing the whole data.
\bigbreak
In the event that a user wants to revoke access to a particular vendor,\ he could do so by changing the encryption key or by changing the randomness used to generate the access token.\ We suggest that users do not regenerate encryption keys but Kratos will provide users an option to choose between the two.\ As we describe school data of K12 students in principle it is the school or school district administrator who controls such access.\ The difference is that the system will further establish accountability and auditability to students with the capacity to gain further control over their data when they become legally independent.
\bigbreak
Furthermore,\ since schools have their own sets of policies,\ Kratos does not strictly enforce a set of practices for users (at school or district level) to follow.\ This ensures that adoption of Kratos is not constrained by a certain set of rules.\ At the same time,\ Kratos defines a set of minimum requirements to be on board to ensure good practices on data protection are followed.
\bigbreak
Kratos also does not enforce how commitments need to be generated and stored,\ leaving it as an option for schools to provide their feedback.\ The system we propose would include a centralized time-stamping server,\ a permissioned blockchain with the different schools as the nodes,\ and simple time-stamping commitments to an existing blockchain.\ All three have their benefits and constraints and Kratos would enable schools to customize this to their requirements.
\bigbreak

>>>>

\textit{Abstract}
Growing digitization has made data ownership an important focus point for institutions and students.\ Broadly,\ there are three issues,\ which require urgent attention for optimization of data privacy,\ literacy,\ and utilization.\ First,\ schools globally are equivocal about data generated by and about students as a result of the digitization of instruction,\ learning,\ and assessment.\ They lack necessary frameworks for data literacy,\ data interoperability,\ and optimization,\ while maintaining privacy and control.\ Second,\ the scale,\ source,\ and nature of school data makes its interoperability impractical,\ resulting in an inability to assess the true impact of educational technologies on instruction and learning.\ Third,\ while data helps teachers improve pedagogical practices,\ an increasingly data-driven decision-making process suggests that student dimensions of learning and equitable participation in curriculum design becomes secondary.\ Finding a balance between data-driven decision making and student voice is critical for an efficient school ecosystem.\ In this paper,\ we introduce Kratos: an immutable decentralised data management system that provides data privacy and applied data literacy while empowering students with a user interface for data governance and active participation in the school ecosystem.\ Using the advantages of blockchain technologies,\ Kratos enables easy authentication and access to data.\ The objective of Kratos is thus to equip students and schools with the ability to access,\ manage, and control their data and to understand how,\ why,\ and by whom data is accessed without compromising student agency and privacy.\ This paper describes proof of concept for Kratos, its benefits to students and schools, and necessary future work.
\bigbreak
\newpage
\section{Introduction}

To improve work,\ school practitioners need data interoperability and educational information from data mining (Bienkowski,\ Feng,\ and Means,\ 2012; Snyder et al.,\ 2016).\ Broadly, this need emanates from the requirement to measure school effectiveness by combining student achievement,\ data,\ and accountability (Militello et al.,\ 2013).

@@

\subsection{Data control}
In conventional systems,\ data control or 'ownership' can be proved with the help of an access token but this provides no guarantee on when the owner came into possession of the data.\ In order to attest 'ownership' at a specific point in time,\ we need time-stamping services like those described by Todd (2016),\ Szalachowski (2018),\ and Massias et al.(1999).\ A time-stamping service requires something to be committed along with the time-stamp and Kratos envisions this to be a cryptographic hash which also acts as an access token which vendors can use to access any student-generated data.
\bigbreak
The aim of using a cryptographic hash like SHA-3 (Bertoni et al.,\ 2012) is to ensure a uniquely random reference to the data to avoid out-of-channel data leaks.\ Kratos suggests using an element of randomness like a salt to generate the hash in order to have the ability to revoke tokens by regenerating randomness.\ Kratos enforces that all schools encrypt their data before creating access tokens to mitigate the risk of loss/theft of data.\ Past research (Zetter,\ 2016) shows that firms are willing to circumnavigate laws to collect data and encrypting data by design ensures that no third party can have access without being granted so explicitly.
\bigbreak
Kratos suggests that encryption be done at the student level but this can also be at the school level depending on existing frameworks and rules surrounding the school.\ If the school chooses to encrypt student data on behalf of the student,\ Kratos enforces that the school use a unique key for each individual to minimize the risk of key theft.\ In addition to this,\ Kratos suggests that schools store their encrypted data in a distributed file storage system like IPFS to ensure data redundancy and availability in case of a setback.\ Storing data on IPFS also makes it easier to create time-stamps since it is sufficient to reference the IPFS pointer instead of potentially hashing the whole data.\ It is to be noted that schools would not be burdened with the load of encrypting data and choosing between different levels of encryption would be a simple click on Kratos' interface.
\bigbreak
In the event that a user wants to revoke access to a particular vendor,\ he could do so by changing the encryption key or by changing the randomness used to generate the access token.\ We suggest that users do not regenerate encryption keys but Kratos will provide users an option to choose between the two.\ As we describe school data of K12 students in principle it is the school or school district administrator who controls such access.\ The difference is that the system will further establish accountability and auditability to students with the capacity to gain further control over their data when they become legally independent.
\bigbreak
Furthermore,\ since schools have their own sets of policies,\ Kratos does not strictly enforce a set of practices for users (at school or district level) to follow.\ This ensures that adoption of Kratos is not constrained by a certain set of rules.\ At the same time,\ Kratos defines a set of minimum requirements to be on board to ensure good practices on data protection are followed.
\bigbreak
Kratos also does not enforce how commitments need to be generated and stored,\ leaving it as an option for schools to provide their feedback.\ The system we propose would include a centralized time-stamping server,\ a permissioned blockchain with the different schools as the nodes,\ and simple time-stamping commitments to an existing blockchain.\ All three have their benefits and constraints and Kratos would enable schools to customize this to their requirements.
\bigbreak
